% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}

\textbf{Genome Wide Association Study (GWAS)}

\textbf{Student Name:} Prit Desai\\
\textbf{UTA ID:} 1002170533\\
\textbf{Course:} CSE 5370 - Bioinformatics\\
\textbf{Assignment:} Homework 1

\textbf{1.2 Fisher's Exact Test}

\textbf{Null Hypothesis Explanation}

The null hypothesis (H₀) for the Fisher's Exact Test states that there
is no association between the presence of the C-allele at a given SNP
and the complex genetic trait being studied. In other words, the odds
ratio between case and control groups for the C-allele is equal to 1.
Rejecting this null hypothesis suggests that the SNP may be associated
with the trait.

\textbf{Choice of Alternative Hypothesis}

The test uses the alternative="two-sided" option because the analysis
aims to detect any association between SNP presence and the trait,
whether positive or negative. This ensures that SNPs with either
increased or decreased associations are identified.

\textbf{Number of Significant SNPs}

\begin{itemize}
\item
  \textbf{Significant SNPs under Original Threshold (5 × 10⁻⁸):} 11
\item
  \textbf{Significant SNPs under Bonferroni Correction (5 × 10⁻¹¹):} 3
\end{itemize}

\textbf{1.3 Bonferroni Correction}

The Bonferroni correction addresses the multiple hypothesis testing
issue by adjusting the significance threshold. Since 1000 SNPs are
tested, the corrected p-value threshold is:

5×10−81000=5×10−11\textbackslash frac\{5 \textbackslash times
10\^{}\{-8\}\}\{1000\} = 5 \textbackslash times 10\^{}\{-11\}

This correction reduces the likelihood of Type I errors but increases
the chance of Type II errors, making the test more conservative and
resulting in fewer significant SNPs compared to the original threshold.

\textbf{1.4 Manhattan Plot Interpretation}

The Manhattan plot visualizes the -log₁₀(p-values) of each SNP locus.
Higher points indicate stronger evidence against the null hypothesis.

\textbf{Key Observations:}

\begin{itemize}
\item
  \textbf{Red dashed line:} Original significance threshold (5 × 10⁻⁸)
\item
  \textbf{Blue dashed line:} Bonferroni-corrected threshold (5 × 10⁻¹¹)
\item
  \textbf{Points above the lines:} Statistically significant SNPs
\item
  A cluster of significant SNPs between loci 450 and 470 indicates a
  potential genomic region associated with the trait.
\end{itemize}

\textbf{Files Included in Submission}

\begin{itemize}
\item
  1002170533.csv: Generated dataset
\item
  results.csv: SNP names, p-values, and significance indicators
\item
  manhattan\_plot.png: Visualization of GWAS results
\item
  GWAS\_Writeup.pdf: This document
\item
  readme.txt: Instructions to run the code
\item
  Source Code:

  \begin{itemize}
  \item
    datasetGenerator.py: Dataset generation
  \item
    HW1.py: GWAS analysis (Fisher's test and Manhattan plot)
  \end{itemize}
\end{itemize}

\textbf{Instructions to Run Code}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Install Python (version ≥ 3.8) and required libraries: numpy, scipy,
  matplotlib.
\item
  Activate the virtual environment:
\item
  .\textbackslash.venv\textbackslash Scripts\textbackslash Activate
\item
  Run the dataset generator:
\item
  python datasetGenerator.py -\/-ID 1002170533
\item
  Run the analysis script:
\item
  python HW1.py
\item
  Outputs:

  \begin{itemize}
  \item
    results.csv: Contains p-values and significance flags
  \item
    manhattan\_plot.png: Displayed and saved automatically
  \end{itemize}
\end{enumerate}

\textbf{Difficulty \& Time Spent}

\begin{itemize}
\item
  \textbf{Estimated Time:} 8 hours
\item
  \textbf{Challenges:} Understanding the Bonferroni correction and
  interpreting the plot required extra time.
\end{itemize}

\end{document}
